\MakeUppercase{Graph Product Structure for Non-Minor-Closed Classes}

### 2022-03-10 k-planar.tex
Some stuff

@@

Let $G$ be a $1$-plane multigraph with no two parallel edges on the boundary of a single face.  By Euler's Formula, $G$ has $O(|V(G)|)$ vertices.  We further assume that $G$ is edge-maximal, in the following sense:  if two vertices $v$ and $w$ of $G$ appear on a common face $F$, then there is an edge $vw\in E(G)$ that is contained in the boundary of $F$.  See  \cref{one_planar_example}(b) for an example.  Any $1$-plane multigraph $G$ with at least three vertices can be made edge-maximal by the addition of at most $3|V(G)|-6$ edges since each edge added this way has endpoints in $V(G)$ and does not cross any other edge of $G$.

@@

If $W$ contains a dummy vertex $w_i$ then neither $w_{i-1}$ nor $w_{i+1}$ is a dummy vertex since if $w_{i-1}$ (respectively $w_{i+1}$) were a dummy vertex then the edge of $G$ that contains $w_{i-1}w_i$ (respectively $w_iw_{i+1}$) would be involved in at least two crossings.  Therefore each dummy vertex $w_i$ of $W$ is surrounded by two vertices $w_{i-1},w_{i+1}\in V(G)$.  Since $w_i$ has degree $4$ in $G_0$, $w_{i-1}\neq w_{i+1}$.  Since $G$ is edge-maximal the edge $w_{i-1}w_i$ is an edge of $G$ on the boundary of $F$.  Therefore, if $F$ has a dummy vertex $w_i$ in its facial walk $W$, then $W=w_{i-1},w_i,w_{i+1}$ and $F$ is bounded by three edges of $G_0$, each of which is contained in a different edge of $G$.

@@

Consider some crossing $(p,vw,xy)$ in $G$.  The point $p$ at which this crossing occurs is a vertex of $G_0$ that has degree $4$.  There are four faces $F_1,\ldots,F_4$ of $G_0$ that contain $p$.  Each of these four faces of is of Type~(ii), so the boundary of $F_1\cup\cdots\cup F_4$ includes a $4$-cycle $C:=vxwy$ in $G$ whose edges are not involved in any crossings.  We call the subgraph $K:=C\cup\{vw,xy\}$ a \defin{kite} of $G$; see \cref{one_planar_example}(a).  The edges $vw$ and $xy$ are called \defin{spars} of $K$.  The cycle $vxwy$ is called the \defin{sail} of $K$.  Thus any edge that is a spar of a kite $K$ is not part of a sail of any kite $K'$.

@@

  The proof is by induction on the number of inner vertices of $N$, i.e., on $|V(N)\setminus V(F)|$.  If $N$ has no inner vertices, then we can simply take $\PP:=\{P_1,\ldots,P_k\}$ and verify the preconditions of the lemma (the `setup') ensure that $\PP$ satisfies the requirements of the lemma.  Therefore, we may assume that $N$ has at least one inner vertex.

>>>>

\usepackage{todonotes}

@@

Let $G$ be a $1$-plane graph.  
% Then $G$ has $O(V(G)$ edges since the edges of $G$ can be partitioned into two planar graphs by placing one edge from each crossing pair into a second set.
% \footnote{We may assume that $G$ has no parallel edges, since removing all but one copy from each set of parallel edges does not change the set of adjacent pairs of vertices in $G$, so any supergraph that contains}
We make $G$ into an \emph{edge-maximal} $1$-plane multigraph, by repeating the following operation:  If two vertices $v$ and $w$ of $G$ appear on a common face $F$ and there is no edge $vw\in E(G)$ contained in the boundary of $F$, then add $vw$, embedded in the face $F$.  This may introduce parallel edges into $G$, but no two such edges will appear on the boundary of a common face.  
% This increases the number of edges of $G$, by $O(|V(G)|)$ since these newly added edges do not cross any other edges of $G$.

@@

If $W$ contains a dummy vertex $w_i$ then neither $w_{i-1}$ nor $w_{i+1}$ is a dummy vertex since if $w_{i-1}$ (respectively $w_{i+1}$) were a dummy vertex then the edge of $G$ that contains $w_{i-1}w_i$ (respectively $w_iw_{i+1}$) would be involved in at least two crossings.  Therefore each dummy vertex $w_i$ of $W$ is surrounded by two non-dummy vertices $w_{i-1},w_{i+1}\in V(G)$.  Since $w_i$ has degree $4$ in $G_0$, $w_{i-1}\neq w_{i+1}$.  Since $G$ is edge-maximal the edge $w_{i-1}w_i$ is an edge of $G$ on the boundary of $F$.  Therefore, if $F$ has a dummy vertex $w_i$ in its facial walk $W$, then $W=w_{i-1},w_i,w_{i+1}$ and $F$ is bounded by three edges of $G_0$, each of which is contained in a different edge of $G$.

@@

Consider some crossing $(p,vw,xy)$ in $G$.  The point $p$ at which this crossing occurs is a vertex of $G_0$ that has degree $4$.  There are four faces $F_1,\ldots,F_4$ of $G_0$ each with $p$ on their boundary.  Each of these four faces of is of Type~(ii), so the boundary of $F_1\cup\cdots\cup F_4$ includes a $4$-cycle $C:=vxwy$ in $G$ whose edges are not involved in any crossings.  We call the subgraph $K:=C\cup\{vw,xy\}$ a \defin{kite} of $G$; see \cref{one_planar_example}(a).  The edges $vw$ and $xy$ are called \defin{spars} of $K$.  The cycle $vxwy$ is called the \defin{sail} of $K$.  Thus any edge that is a spar of a kite $K$ is not part of a sail of any kite $K'$.\todo{Drop the definitions of kite and spar, we hardly use them.}

@@

  The proof is by induction on the number of inner vertices of $N$, i.e., on $|V(N)\setminus V(F)|$.  If $N$ has no inner vertices, then we can simply take $\PP:=\{P_1,\ldots,P_k\}$ and verify that the preconditions of the lemma (the `setup') ensure that $\PP$ satisfies the requirements of the lemma.  Therefore, we may assume that $N$ has at least one inner vertex.
